---
title: "Project memo"
author: "Y & J Code Craze"
output: html_document
---

This document should contain a detailed account of the data clean up for your data and the design choices you are making for your plots. For instance you will want to document choices you've made that were intentional for your graphic, e.g. color you've chosen for the plot. Think of this document as a code script someone can follow to reproduce the data cleaning steps and graphics in your handout.

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
library(tidyverse)
library(broom)
library(leaflet)
library(colorspace)
library(sf)
library(visdat)
library(naniar)
library(UpSetR)
library(RColorBrewer)

Internship_Data_Untidy <- read.csv("/cloud/project/data/Internship_Data_Untidy.csv")
Tidy_Mapped_Clean <- read.csv("/cloud/project/data/Tidy_Mapped_Clean.csv")
Internship_Leaflet_Official_Complete <- read_csv("/cloud/project/data/Internship_Leaflet_Official - Complete.csv")
```

## Data Clean Up Steps for Overall Data

### Step 1: Tidy the Term Column

#### Step 1 a.) Change Entries with unknown term data to "NA" and unnesesary remove spaces before and after an entry 
```{r clear-unknown-term-data}
Internship_Data_Untidy %>% 
  mutate(Term = case_when( 
    Term %in% c("NA", "Not Stated", "-", "", "NA ", "All Year", "n/a", "Year", "Year-round", "N/A", "Previous Work Experience", "Nine Months", "Prior Fulfillment"))) ~ "NA" %>% 
  mutate(Term = str_squish(Term))
```

#### Step 1 b.) Ensure all terms corespond to one COA term.

Because we are using the last recorded year from the "year" column, we are also using the last recorded term. The last term of each entry should correspond to the year the internship was completed. 

```{r filtering-for-term}
Tidy_Term <- Internship_Data_Untidy %>% 
mutate(Term = case_when(
  Term %in% c("Fall", "Winter, Spring, Fall", "Winter, Fall", "Summer-Fall", "Spring, Fall", "Summer/Fall", "Summer, Fall", "Spring, Summer, Fall", "Summer,Fall", "Winter+Fall") ~ "Fall",
  Term %in% c("Winter", "Winter, Fall, Winter", "Fall/Winter", "December-December", "Summer, Fall, Winter", "Fall-Winter", "Fall, Winter", "Summer-Winter") ~ "Winter",
  Term %in% c("Spring", "Fall, Winter, Spring", "spring", "Winter, Spring", "Winter-Spring", "Novermber-Spring", "Summer, Spring, Spring") ~ "Spring",
  Term %in% c("Summer", "Fall, Summer", "August", "Winter-Summer", "Winter, Spring, Summer", "Winter, Summer", "Spring, Summer", "Fall+Summer", "Spring/Summer", "Fall, Winter, Spring, Summer", "Summe", "Summer 2011") ~ "Summer"
))
```

#### Step 1 c.) Yay it's tidy!
```{r ta-da}
Tidy_Term %>% 
  group_by(Term) %>%
  summarise(count = n())
```

### Step 2: Tidy the Year Column (All work applied to data frame "Tidy_Term")

#### Step 2 a.) Check for all unique variables within the Year coloumn 

```{r year-variable-check}

unique(Tidy_Term$Year)
```

#### Step 2 b.) Ensure all variables follow same format

The format is simply to state the year as a 4 digit integer. "XXXX". 
We were instructed by Jeffry to use the latest year for entries that span across multiple years.

```{r mutating-unique-years}
Tidy_Time <- Tidy_Term %>% 
mutate(Year = case_when(
  Year %in% c("1970") ~ "1970",
  Year %in% c("1973", "1973, 1974, 1975") ~ "1973",
  Year %in% c("1974") ~ "1974",
  Year %in% c("1975", "1974, 1975") ~ "1975",
  Year %in% c("1976") ~"1976",
  Year %in% c("1977", "1977, 1975, 1972") ~ "1977",
  Year %in% c("1978", "1977,1978") ~ "1978",
  Year %in% c("1979","1976-79") ~ "1979",
  Year %in% c("1980", "1978, 1979, 1980", "1979 and 1980") ~ "1980",
  Year %in% c("1981") ~ "1981",
  Year %in% c("1982", "1981, 1982") ~ "1982",
  Year %in% c("1983","1981-1983") ~ "1983",
  Year %in% c("1984") ~ "1984",
  Year %in% c("1985", "1984, 1985") ~ "1985",
  Year %in% c("1986", "1985, 1986") ~ "1986",
  Year %in% c("1987") ~"1987",
  Year %in% c("1988", "1987, 1988") ~ "1988",
  Year %in% c("1989") ~ "1989",
  Year %in% c("1990") ~ "1990",
  Year %in% c("1991") ~ "1991",
  Year %in% c("1992") ~ "1992",
  Year %in% c("1993", "1987-1993") ~ "1993",
  Year %in% c("1994") ~ "1994",
  Year %in% c("1995") ~ "1995",
  Year %in% c("1996", "1994, 1995, 1996", "1995, 1996") ~ "1996",
  Year %in% c("1997") ~ "1997",
  Year %in% c("1998", "	1997, 1998") ~ "1998",
  Year %in% c("1999", "1997, 1998, 1999", "1998/1999") ~ "1999",
  Year %in% c("2000") ~ "2000",
  Year %in% c("2001") ~ "2001",
  Year %in% c("2002", "2001, 2002") ~ "2002",
  Year %in% c("2003") ~ "2003",
  Year %in% c("2004") ~"2004",
  Year %in% c("2005", "2003, 2004, 2005") ~ "2005",
  Year %in% c("2006") ~ "2006",
  Year %in% c("2007", "2006, 2007") ~ "2007",
  Year %in% c("2008", "1996-2008") ~ "2008",
  Year %in% c("2009","2002 - 2009") ~ "2009",
  Year %in% c("2010", "2009-2010") ~ "2010",
  Year %in% c("2011","	2009-2011") ~ "2011",
  Year %in% c("2012") ~ "2012",
  Year %in% c("2013") ~ "2013",
  Year %in% c("2014", "	2013-2014") ~ "2014",
  Year %in% c("2015", "2014-15") ~ "2015",
  Year %in% c("2016", "	2015, 2016") ~ "2016",
  Year %in% c("2017", "	2015, 2016, 2017") ~ "2017",
  Year %in% c("2018") ~ "2018",
  Year %in% c("2019") ~ "2019",
  Year %in% c("2020") ~ "2020",
  Year %in% c("2021") ~ "2021",
  Year %in% c("2022") ~ "2022",
  Year %in% c("2023") ~ "2023"
))
```

#### Step 2 c.) Double check unique terms

```{r year-variable-recheck}
Tidy_Time %>% 
  group_by(Year) %>%
  summarise(count = n())
```
Yay! it's all tidy! 

### Step 3 Create a Decade Variable 
We thought that it would be easier to visualize change over time if we used decades instead of years. We also wanted to ensure that our plots were readable. Because this data set is growing, there is not very much information from the 2020s. Internships have not decreased on the last 4 years (as far as we know), this is just what is availible in our dataset. 

```{r create-decade-variable}
Internship_Leaflet_Official_Complete <- Internship_Leaflet_Official_Complete %>% 
  mutate(Decade = (floor(Year/10)*10))
```

### Step 4 Create Local, Domestic, and International Catagories 
We wanted to compare local, domestic, and international internships across multiple decades.
Most internships were formatted "City, US state" or "City, Country." We went into excel and changed entries to match this format. Canadian state names were changed to "City, Canada." Entries that were formatted "City, Non-US State, Country" were changed to "City, Country." We then separated the state or country at the comma. Anything with a US state after the comma was labeled "Domestic" and any country after the comma was labeled "International. 
All internships located in towns on MDI were counted as local. This was the easiest way to categorize local internships rather than having to decide which off-island towns/areas counted as local.
```{r create-location-type}
us_states <- c("Alabama", "Alaska", "Arizona", "Arkansas", "California", "Colorado", 
               "Connecticut", "Delaware", "Florida", "Georgia", "Hawaii", "Idaho", 
               "Illinois", "Indiana", "Iowa", "Kansas", "Kentucky", "Louisiana", 
               "Maine", "Maryland", "Massachusetts", "Michigan", "Minnesota", 
               "Mississippi", "Missouri", "Montana", "Nebraska", "Nevada", 
               "New Hampshire", "New Jersey", "New Mexico", "New York", 
               "North Carolina", "North Dakota", "Ohio", "Oklahoma", "Oregon", 
               "Pennsylvania", "Rhode Island", "South Carolina", "South Dakota", 
               "Tennessee", "Texas", "Utah", "Vermont", "Virginia", "Washington", 
               "West Virginia", "Wisconsin", "Wyoming")
Internship_Leaflet_Official_Complete <- Internship_Leaflet_Official_Complete %>%
mutate(region = str_extract(Location, "(?<=, ).+$"),
       Location_Type = case_when(
         Location %in% c("Bar Harbor, Maine", "Northeast Harbor, Maine", "Southwest Harbor, Maine", "Otter Creek, Maine", "Seal Cove, Maine", "Mount Desert, Maine", "Mount Desert Island, Maine", "Hulls Cove, Maine", "Town Hill, Maine", "Somesville, Maine", "Salsbury Cove, Maine") ~ "Local",
      region %in% us_states ~ "Domestic",
      TRUE ~ "International"))
```


### ggsave example for saving plots

```{r starwars ggsave example}
exp1 <- starwars |>
  filter(mass < 1000, 
         species %in% c("Human", "Cerean", "Pau'an", "Droid", "Gungan")) |>
  ggplot() +
  geom_point(aes(x = mass, 
                 y = height, 
                 color = species)) +
  labs(x = "Weight (kg)", 
       y = "Height (m)",
       color = "Species",
       title = "Weight and Height of Select Starwars Species",
       caption = paste("This data comes from the starwars api: https://swapi.py43.com"))


ggsave("example-starwars.png", width = 4, height = 4)

ggsave("example-starwars-wide.png", width = 6, height = 4)
```




```{r test-plot-3}
#building leaflet
#Things to change: decade, clustering (for later), sliding scale
pal <- colorFactor(palette = c("red", "orange", "yellow", "green", "blue", "purple"), domain = Tidy_Mapped_Clean$Decade)

leaflet(Tidy_Mapped_Clean) %>%
  addTiles() %>%
  setView(lng = mean(Tidy_Mapped_Clean$Longitude, na.rm = TRUE), 
          lat = mean(Tidy_Mapped_Clean$Latitude, na.rm = TRUE), 
          zoom = 2) %>%
  addCircleMarkers(data = Tidy_Mapped_Clean %>% filter(Decade == "1970"),
                   ~Longitude, 
                   ~Latitude, 
                   color = ~pal(Decade),
                   group = "1970s",
                   popup = ~paste(Internship.Site, "<br>", Location),
                   clusterOptions = markerClusterOptions()) %>% 
  addCircleMarkers(data = Tidy_Mapped_Clean %>% filter(Decade == "1980"),
                   ~Longitude, 
                   ~Latitude, 
                   color = ~pal(Decade),
                   group = "1980s",
                   popup = ~paste(Internship.Site, "<br>", Location),
                   clusterOptions = markerClusterOptions()) %>% 
  addCircleMarkers(data = Tidy_Mapped_Clean %>% filter(Decade == "1990"),
                   ~Longitude, 
                   ~Latitude, 
                   color = ~pal(Decade),
                   group = "1990s",
                   popup = ~paste(Internship.Site, "<br>", Location),
                   clusterOptions = markerClusterOptions()) %>% 
  addCircleMarkers(data = Tidy_Mapped_Clean %>% filter(Decade == "2000"),
                   ~Longitude, 
                   ~Latitude, 
                   color = ~pal(Decade),
                   group = "2000s",
                   popup = ~paste(Internship.Site, "<br>", Location),
                   clusterOptions = markerClusterOptions()) %>% 
  addCircleMarkers(data = Tidy_Mapped_Clean %>% filter(Decade == "2010"),
                   ~Longitude, 
                   ~Latitude, 
                   color = ~pal(Decade),
                   group = "2010s",
                   popup = ~paste(Internship.Site, "<br>", Location),
                   clusterOptions = markerClusterOptions()) %>% 
  addCircleMarkers(data = Tidy_Mapped_Clean %>% filter(Decade == "2020"),
                   ~Longitude, 
                   ~Latitude, 
                   color = ~pal(Decade),
                   group = "2020s",
                   popup = ~paste(Internship.Site, "<br>", Location),
                   clusterOptions = markerClusterOptions()) %>% 
  addLayersControl(overlayGroups = c("1970s", "1980s", "1990s", "2000s", "2010s", "2020s"),
                   options = layersControlOptions())
```

```{r test-plot-4}
leaflet(Tidy_Mapped_Clean) %>%
  addTiles() %>%
  setView(lng = mean(Tidy_Mapped_Clean$Longitude, na.rm = TRUE), 
          lat = mean(Tidy_Mapped_Clean$Latitude, na.rm = TRUE), 
          zoom = 2) %>%
  addMarkers(~Longitude, ~Latitude,
             clusterOptions = markerClusterOptions(),
             popup = ~paste(Internship.Site, "<br>", Location))
```

```{r test-plot-5}
pal <- colorFactor(palette = c("red", "orange", "yellow", "green", "blue", "purple"), domain = Internship_Leaflet_Official_Complete$Decade)

leaflet(Internship_Leaflet_Official_Complete) %>%
  addTiles() %>%
  setView(lng = mean(Internship_Leaflet_Official_Complete$Longitude, na.rm = TRUE), 
          lat = mean(Internship_Leaflet_Official_Complete$Latitude, na.rm = TRUE), 
          zoom = 2) %>%
  addCircleMarkers(data = Internship_Leaflet_Official_Complete %>% filter(Decade == "1970"),
                   ~Longitude, 
                   ~Latitude, 
                   color = ~pal(Decade),
                   group = "1970s",
                   popup = ~paste(Internship.Site, "<br>", Location, "<br>", Year)) %>% 
  addCircleMarkers(data = Internship_Leaflet_Official_Complete %>% filter(Decade == "1980"),
                   ~Longitude, 
                   ~Latitude, 
                   color = ~pal(Decade),
                   group = "1980s",
                   popup = ~paste(Internship.Site, "<br>", Location, "<br>", Year)) %>% 
  addCircleMarkers(data = Internship_Leaflet_Official_Complete %>% filter(Decade == "1990"),
                   ~Longitude, 
                   ~Latitude, 
                   color = ~pal(Decade),
                   group = "1990s",
                   popup = ~paste(Internship.Site, "<br>", Location, "<br>", Year)) %>% 
  addCircleMarkers(data = Internship_Leaflet_Official_Complete %>% filter(Decade == "2000"),
                   ~Longitude, 
                   ~Latitude, 
                   color = ~pal(Decade),
                   group = "2000s",
                   popup = ~paste(Internship.Site, "<br>", Location, "<br>", Year)) %>% 
  addCircleMarkers(data = Internship_Leaflet_Official_Complete %>% filter(Decade == "2010"),
                   ~Longitude, 
                   ~Latitude, 
                   color = ~pal(Decade),
                   group = "2010s",
                   popup = ~paste(Internship.Site, "<br>", Location, "<br>", Year)) %>% 
  addCircleMarkers(data = Internship_Leaflet_Official_Complete %>% filter(Decade == "2020"),
                   ~Longitude, 
                   ~Latitude, 
                   color = ~pal(Decade),
                   group = "2020s",
                   popup = ~paste(Internship.Site, "<br>", Location, "<br>", Year)) %>% 
  addLayersControl(overlayGroups = c("1970s", "1980s", "1990s", "2000s", "2010s", "2020s")) %>%
  addLegend("bottomright", 
            pal = pal, 
            values = Internship_Leaflet_Official_Complete$Decade[!is.na(Internship_Leaflet_Official_Complete$Decade)],
            title = "Decade",
            labFormat = labelFormat(prefix = ""))
```

```{r tes-plot-6}
leaflet(Internship_Leaflet_Official_Complete) %>%
  addTiles() %>%
  setView(lng = mean(Internship_Leaflet_Official_Complete$Longitude, na.rm = TRUE), 
          lat = mean(Internship_Leaflet_Official_Complete$Latitude, na.rm = TRUE), 
          zoom = 2) %>%
  addMarkers(~Longitude, ~Latitude, 
             clusterOptions = markerClusterOptions(),
             popup = ~paste(Internship.Site, "<br>", Location, "<br>", Year))
```

```{r test-plot-lucky-number-7}
Internship_Leaflet_Official_Complete %>% 
  filter(Decade != "NA") %>% 
  ggplot(aes(x = Location_Type,
             fill = Location_Type)) +
 geom_bar() +
   facet_wrap(~Decade) +
  scale_fill_viridis_d()
```

```{r test-plot-number-8}
Internship_Leaflet_Official_Complete %>% 
  ggplot(aes(y = Year,
         fill = Location_Type)) + 
  geom_bar(position = "fill") + 
  scale_fill_viridis_d()
```



```{r plot-1-final}
Internship_Leaflet_Official_Complete %>% 
  drop_na(Term) %>% 
  drop_na(Decade) %>% 
  group_by(Decade, Term) %>%
  summarise(count = n()) %>%
  ungroup() %>% 
  mutate(Term = fct_relevel(Term, "Fall", "Winter", "Spring", "Summer")) %>%  
ggplot(aes(x = Decade, y = count, fill = as.numeric(Decade))) + 
  geom_bar(stat = "identity", position = "stack") + 
  facet_wrap(~Term, nrow = 4, scales = "free_y") + 
  theme_minimal() +  
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        legend.key.size = unit(0.7, "lines")) +
  scale_fill_viridis_c(direction = -1) +
  coord_flip() +
  theme(legend.position = "top") + 
  guides(fill = guide_legend(nrow = 1, byrow = TRUE)) +
  labs(
    y = "Number of Internships", 
    title = "Student Internship Year Breakup by Term",
    subtitle = "Data from College of the Atlantic (1970 to 2023)",
    fill = "Decade")
```

### Plot 2: Leaflet Map of Internships by Decade

```{r plot-2-final}
Top_10_Sites <- Internship_Leaflet_Official_Complete %>% 
  filter(!is.na(Internship.Site) & Internship.Site != "Prior Fulfillment of Internship Requirement") %>% 
  mutate(Internship.Site = case_when( 
    Internship.Site %in% c("MDI Biological Laboratory", "Mount Desert Island Biological Laboratory") ~ "MDI Biological Laboratory",
    TRUE ~ Internship.Site
  )) %>% 
  mutate(Internship.Site = case_when( 
    Internship.Site %in% c("The Jackson Laboratory", "Jackson Laboratory") ~ "Jackson Laboratory",
    TRUE ~ Internship.Site
  )) %>% 
  mutate(Internship.Site = str_squish(Internship.Site)) %>% 
mutate(Internship.Site = case_when( 
    Internship.Site %in% c("YMCA Camp Jordan", "YMCA Camp Erdman", "YMCA", "Mount Desert Island YMCA", "MDI YMCA", "Bar Harbor YMCA", "Bangor YMCA Camp Jordan") ~ "YMCA",
    TRUE ~ Internship.Site
  )) %>% 
  mutate(Internship.Site = str_squish(Internship.Site)) %>% 
mutate(Internship.Site = case_when( 
    Internship.Site %in% c("Summer Field Studies College of the Atlantic", "Summer Field Studies at COA", "Summer Field Studies") ~ "Summer Field Studies",
    TRUE ~ Internship.Site
  )) %>% 
  mutate(Internship.Site = str_squish(Internship.Site)) %>%
mutate(Internship.Site = case_when( 
    Internship.Site %in% c("Mount Desert Island Hospital", "MDI Hospital") ~ "Mount Desert Island Hospital",
    TRUE ~ Internship.Site
  )) %>% 
  mutate(Internship.Site = str_squish(Internship.Site)) %>%
mutate(Internship.Site = case_when( 
    Internship.Site %in% c("Island Reaserch Center", "Alice Eno Field Station on Great Duck Island", "Alice Eno Research Center", "Great Duck Island", "Alice Eno Biological Station and Mount Desert Rock") ~ "Great Duck Island Research Center",
    TRUE ~ Internship.Site
  )) %>% 
  mutate(Internship.Site = str_squish(Internship.Site)) %>%
mutate(Internship.Site = case_when( 
    Internship.Site %in% c("Acadia National Park Resource Management", "Acadia National Park") ~ "Acadia National Park",
    TRUE ~ Internship.Site
  )) %>% 
  mutate(Internship.Site = str_squish(Internship.Site)) %>%
  count(Internship.Site) %>% 
  arrange(desc(n)) %>% 
  slice_head(n = 10)

Top_10_Sites %>% 
  ggplot(aes(x = reorder(Internship.Site, n), y = n, fill = reorder(Internship.Site, n))) +
  geom_bar(stat = "identity") + 
  theme_minimal() +
  theme(legend.position = "none") +
  coord_flip() + 
  scale_fill_viridis_d(direction = -1)
```

### Plot 3:

```{r test-plot-9}
Internship_Leaflet_Official_Complete %>%
  count(Year, Location_Type) %>%
  ggplot(aes(x = Year, y = n, colour = Location_Type)) +
  geom_line(size = 0.8) +
  theme_minimal() +
  labs(x = "Year", y = "Count", title = "Counts of Internship Locations by Year and Type") + 
  scale_colour_viridis_d()
```

