---
title: "Test Plots"
author: "Y & J Code Craze"
output: html_document
---

This document should contains all of our test plots and the tidying that went along with them. We wanted to use this document to show our process and to domcument the rest of the work we did. 

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
library(tidyverse)
library(broom)
library(leaflet)
library(colorspace)
library(sf)
library(visdat)
library(naniar)
library(UpSetR)
library(RColorBrewer)

Internship_Data_Untidy <- read.csv("/cloud/project/data/Internship_Data_Untidy.csv")
Tidy_Mapped_Clean <- read.csv("/cloud/project/data/Tidy_Mapped_Clean.csv")
Internship_Leaflet_Official_Complete <- read_csv("/cloud/project/data/Internship_Leaflet_Official - Complete.csv")
```

## Data Clean Up + Extra Steps Associated with Different Test Plots
A lot of this is repeat info - sorry. 

```{r clear-unknown-term-data}
Internship_Data_Untidy %>% 
  mutate(Term = case_when( 
    Term %in% c("NA", "Not Stated", "-", "", "NA ", "All Year", "n/a", "Year", "Year-round", "N/A", "Previous Work Experience", "Nine Months", "Prior Fulfillment"))) ~ "NA" %>% 
  mutate(Term = str_squish(Term))
```

#### Step 1 b.) Ensure all terms corespond to one COA term.

```{r filtering-for-term}
Tidy_Term <- Internship_Data_Untidy %>% 
mutate(Term = case_when(
  Term %in% c("Fall", "Winter, Spring, Fall", "Winter, Fall", "Summer-Fall", "Spring, Fall", "Summer/Fall", "Summer, Fall", "Spring, Summer, Fall", "Summer,Fall", "Winter+Fall") ~ "Fall",
  Term %in% c("Winter", "Winter, Fall, Winter", "Fall/Winter", "December-December", "Summer, Fall, Winter", "Fall-Winter", "Fall, Winter", "Summer-Winter") ~ "Winter",
  Term %in% c("Spring", "Fall, Winter, Spring", "spring", "Winter, Spring", "Winter-Spring", "Novermber-Spring", "Summer, Spring, Spring") ~ "Spring",
  Term %in% c("Summer", "Fall, Summer", "August", "Winter-Summer", "Winter, Spring, Summer", "Winter, Summer", "Spring, Summer", "Fall+Summer", "Spring/Summer", "Fall, Winter, Spring, Summer", "Summe", "Summer 2011") ~ "Summer"
))
```

#### Step 1 c.) Yay it's tidy!
```{r ta-da}
Tidy_Term %>% 
  group_by(Term) %>%
  summarise(count = n())
```

### Step 2: Tidy the Year Column (All work applied to data frame "Tidy_Term")

#### Step 2 a.) Check for all unique variables within the Year coloumn 

```{r year-variable-check}

unique(Tidy_Term$Year)
```

#### Step 2 b.) Ensure all variables follow same format

```{r mutating-unique-years}
Tidy_Time <- Tidy_Term %>% 
mutate(Year = case_when(
  Year %in% c("1970") ~ "1970",
  Year %in% c("1973", "1973, 1974, 1975") ~ "1973",
  Year %in% c("1974") ~ "1974",
  Year %in% c("1975", "1974, 1975") ~ "1975",
  Year %in% c("1976") ~"1976",
  Year %in% c("1977", "1977, 1975, 1972") ~ "1977",
  Year %in% c("1978", "1977,1978") ~ "1978",
  Year %in% c("1979","1976-79") ~ "1979",
  Year %in% c("1980", "1978, 1979, 1980", "1979 and 1980") ~ "1980",
  Year %in% c("1981") ~ "1981",
  Year %in% c("1982", "1981, 1982") ~ "1982",
  Year %in% c("1983","1981-1983") ~ "1983",
  Year %in% c("1984") ~ "1984",
  Year %in% c("1985", "1984, 1985") ~ "1985",
  Year %in% c("1986", "1985, 1986") ~ "1986",
  Year %in% c("1987") ~"1987",
  Year %in% c("1988", "1987, 1988") ~ "1988",
  Year %in% c("1989") ~ "1989",
  Year %in% c("1990") ~ "1990",
  Year %in% c("1991") ~ "1991",
  Year %in% c("1992") ~ "1992",
  Year %in% c("1993", "1987-1993") ~ "1993",
  Year %in% c("1994") ~ "1994",
  Year %in% c("1995") ~ "1995",
  Year %in% c("1996", "1994, 1995, 1996", "1995, 1996") ~ "1996",
  Year %in% c("1997") ~ "1997",
  Year %in% c("1998", "	1997, 1998") ~ "1998",
  Year %in% c("1999", "1997, 1998, 1999", "1998/1999") ~ "1999",
  Year %in% c("2000") ~ "2000",
  Year %in% c("2001") ~ "2001",
  Year %in% c("2002", "2001, 2002") ~ "2002",
  Year %in% c("2003") ~ "2003",
  Year %in% c("2004") ~"2004",
  Year %in% c("2005", "2003, 2004, 2005") ~ "2005",
  Year %in% c("2006") ~ "2006",
  Year %in% c("2007", "2006, 2007") ~ "2007",
  Year %in% c("2008", "1996-2008") ~ "2008",
  Year %in% c("2009","2002 - 2009") ~ "2009",
  Year %in% c("2010", "2009-2010") ~ "2010",
  Year %in% c("2011","	2009-2011") ~ "2011",
  Year %in% c("2012") ~ "2012",
  Year %in% c("2013") ~ "2013",
  Year %in% c("2014", "	2013-2014") ~ "2014",
  Year %in% c("2015", "2014-15") ~ "2015",
  Year %in% c("2016", "	2015, 2016") ~ "2016",
  Year %in% c("2017", "	2015, 2016, 2017") ~ "2017",
  Year %in% c("2018") ~ "2018",
  Year %in% c("2019") ~ "2019",
  Year %in% c("2020") ~ "2020",
  Year %in% c("2021") ~ "2021",
  Year %in% c("2022") ~ "2022",
  Year %in% c("2023") ~ "2023"
))
```

#### Step 2 c.) Double check unique terms

```{r year-variable-recheck}
Tidy_Time %>% 
  group_by(Year) %>%
  summarise(count = n())
```
Yay! it's all tidy! 

### Step 3 Create a Decade Variable 
Decade variable for different data sets used to create different plots. 

```{r create-decade-variable-1}
Tidy_Mapped_Clean <- Tidy_Mapped_Clean %>% 
  mutate(Decade = (floor(Year/10)*10))
```
 
```{r create-decade-variable-2}
Internship_Leaflet_Official_Complete <- Internship_Leaflet_Official_Complete %>% 
  mutate(Decade = (floor(Year/10)*10))
```

### Step 4 Create Local, Domestic, and International Catagories 

```{r create-location-type}
us_states <- c("Alabama", "Alaska", "Arizona", "Arkansas", "California", "Colorado", 
               "Connecticut", "Delaware", "Florida", "Georgia", "Hawaii", "Idaho", 
               "Illinois", "Indiana", "Iowa", "Kansas", "Kentucky", "Louisiana", 
               "Maine", "Maryland", "Massachusetts", "Michigan", "Minnesota", 
               "Mississippi", "Missouri", "Montana", "Nebraska", "Nevada", 
               "New Hampshire", "New Jersey", "New Mexico", "New York", 
               "North Carolina", "North Dakota", "Ohio", "Oklahoma", "Oregon", 
               "Pennsylvania", "Rhode Island", "South Carolina", "South Dakota", 
               "Tennessee", "Texas", "Utah", "Vermont", "Virginia", "Washington", 
               "West Virginia", "Wisconsin", "Wyoming")
Internship_Leaflet_Official_Complete <- Internship_Leaflet_Official_Complete %>%
mutate(region = str_extract(Location, "(?<=, ).+$"),
       Location_Type = case_when(
         Location %in% c("Bar Harbor, Maine", "Northeast Harbor, Maine", "Southwest Harbor, Maine", "Otter Creek, Maine", "Seal Cove, Maine", "Mount Desert, Maine", "Mount Desert Island, Maine", "Hulls Cove, Maine", "Town Hill, Maine", "Somesville, Maine", "Salsbury Cove, Maine") ~ "Local",
      region %in% us_states ~ "Domestic",
      TRUE ~ "International"))
```

### Test Plots: Year-Term Bar Polar Charts

Polar chart of terms by year: 

```{r test-plot-1}
Tidy_Time_Summary <- Tidy_Time %>%
  drop_na(Term) %>% 
  drop_na(Year) %>% 
  group_by(Year, Term) %>%
  summarise(count = n()) %>%
  ungroup()

Tidy_Time_Summary %>% 
  ggplot(aes(x = Year, y = count, fill = fct_relevel(Term, c("Summer", "Spring", "Winter", "Fall")))) + 
  geom_bar(stat = "identity", position = "stack") + 
  coord_polar(start = 0) + 
  labs(
    title = "Student Internship Term Beakup by Year",
    subtitle = "Data from College of the Atlantic (1970 to 2023)",
    x = "Year",
    fill = "Term") +
  theme_minimal() +
  theme(axis.title.y = element_blank(),
        axis.line = element_blank(), 
        axis.ticks = element_blank(),
        axis.text.y = element_blank())
```

Polar chart of terms from all years: 

```{r test-plot-2}
Tidy_Time_Summary %>% 
  ggplot(aes(x = Term, y = count, fill = as.numeric(Year))) + 
  geom_bar(stat = "identity", position = "stack") + 
  coord_polar(start = 0) + 
  labs(
    title = "Student Internship Year Breakup by Term",
    subtitle = "Data from College of the Atlantic (1970 to 2023)",
    x = "Term",
    fill = "Year") + 
  theme_minimal() +  
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        legend.key.size = unit(0.7, "lines")) +
  scale_fill_viridis_c(direction = -1)
```

Same idea, but split up by decade instead of year: 

```{r test-plot-3}
Internship_Leaflet_Official_Complete %>% 
  drop_na(Term) %>% 
  drop_na(Decade) %>% 
  group_by(Decade, Term) %>%
  summarise(count = n()) %>%
  ungroup() %>% 
  ggplot(aes(x = Term, y = count, fill = as.numeric(Decade))) + 
  geom_bar(stat = "identity", position = "stack") + 
  coord_polar(start = 0) + 
  labs(
    title = "Student Internship Year Breakup by Term",
    subtitle = "Data from College of the Atlantic (1970 to 2023)",
    x = "Term",
    fill = "Year") + 
  theme_minimal() +  
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        legend.key.size = unit(0.7, "lines")) +
  scale_fill_viridis_c(direction = -1)
```

More variations, but make it a bar plot: 

```{r test-plot-4}
Internship_Leaflet_Official_Complete %>% 
  drop_na(Term) %>% 
  drop_na(Decade) %>% 
  group_by(Decade, Term) %>%
  summarise(count = n()) %>%
  ungroup() %>% 
  ggplot(aes(x = Term, y = count, fill = as.numeric(Decade))) + 
  geom_bar(stat = "identity", position = "stack") + 
  labs(
    title = "Student Internship Year Breakup by Term",
    subtitle = "Data from College of the Atlantic (1970 to 2023)",
    x = "Term",
    fill = "Year") + 
  theme_minimal() +  
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        legend.key.size = unit(0.7, "lines")) +
  scale_fill_viridis_c(direction = -1)
```

```{r test-plot-5}
Internship_Leaflet_Official_Complete %>% 
  drop_na(Term) %>% 
  drop_na(Decade) %>% 
  group_by(Decade, Term) %>%
  summarise(count = n()) %>%
  ungroup() %>% 
ggplot(aes(x = Term, y = count, fill = as.numeric(Decade))) + 
  geom_bar(stat = "identity", position = "fill") + 
  labs(
    title = "Student Internship Year Breakup by Term",
    subtitle = "Data from College of the Atlantic (1970 to 2023)",
    x = "Term",
    fill = "Year") + 
  theme_minimal() +  
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        legend.key.size = unit(0.7, "lines")) +
  scale_fill_viridis_c(direction = -1)
```

### Test Plots: Decade-Location Type Plots

```{r test-plot-lucky-number-7}
Internship_Leaflet_Official_Complete %>% 
  filter(Decade != "NA") %>% 
  ggplot(aes(x = Location_Type,
             fill = Location_Type)) +
 geom_bar() +
   facet_wrap(~Decade) +
  scale_fill_viridis_d()
```
This one is interesting to look at, but there is too much going on. 

```{r test-plot-number-8}
Internship_Leaflet_Official_Complete %>% 
  ggplot(aes(y = Year,
         fill = Location_Type)) + 
  geom_bar(position = "fill") + 
  scale_fill_viridis_d()
```

This one almost made it onto our poster, but we figured that it would be more interesting to have more variation in our visualizations (we have so many bar charts). 

```{r test-plot-3}
Internship_Leaflet_Official_Complete %>%
ggplot(aes(x = Decade,
             fill = Location_Type)) +
 geom_bar(position = "stack") +
  facet_wrap(~Location_Type, nrow = 3, scales = "free_y") +
  scale_fill_viridis_d()
```

### Test Plots: Old Leaflet Map

All internships without all of the lat/lons entered:

```{r test-plot-4}
leaflet(Tidy_Mapped_Clean) %>%
  addTiles() %>%
  setView(lng = mean(Tidy_Mapped_Clean$Longitude, na.rm = TRUE), 
          lat = mean(Tidy_Mapped_Clean$Latitude, na.rm = TRUE), 
          zoom = 2) %>%
  addMarkers(~Longitude, ~Latitude,
             clusterOptions = markerClusterOptions(),
             popup = ~paste(Internship.Site, "<br>", Location))
```

Same thing, but with the decade layers: 

```{r test-plot-3}
#building leaflet
#Things to change: decade, clustering (for later), sliding scale
pal <- colorFactor(palette = c("red", "orange", "yellow", "green", "blue", "purple"), domain = Tidy_Mapped_Clean$Decade)

leaflet(Tidy_Mapped_Clean) %>%
  addTiles() %>%
  setView(lng = mean(Tidy_Mapped_Clean$Longitude, na.rm = TRUE), 
          lat = mean(Tidy_Mapped_Clean$Latitude, na.rm = TRUE), 
          zoom = 2) %>%
  addCircleMarkers(data = Tidy_Mapped_Clean %>% filter(Decade == "1970"),
                   ~Longitude, 
                   ~Latitude, 
                   color = ~pal(Decade),
                   group = "1970s",
                   popup = ~paste(Internship.Site, "<br>", Location),
                   clusterOptions = markerClusterOptions()) %>% 
  addCircleMarkers(data = Tidy_Mapped_Clean %>% filter(Decade == "1980"),
                   ~Longitude, 
                   ~Latitude, 
                   color = ~pal(Decade),
                   group = "1980s",
                   popup = ~paste(Internship.Site, "<br>", Location),
                   clusterOptions = markerClusterOptions()) %>% 
  addCircleMarkers(data = Tidy_Mapped_Clean %>% filter(Decade == "1990"),
                   ~Longitude, 
                   ~Latitude, 
                   color = ~pal(Decade),
                   group = "1990s",
                   popup = ~paste(Internship.Site, "<br>", Location),
                   clusterOptions = markerClusterOptions()) %>% 
  addCircleMarkers(data = Tidy_Mapped_Clean %>% filter(Decade == "2000"),
                   ~Longitude, 
                   ~Latitude, 
                   color = ~pal(Decade),
                   group = "2000s",
                   popup = ~paste(Internship.Site, "<br>", Location),
                   clusterOptions = markerClusterOptions()) %>% 
  addCircleMarkers(data = Tidy_Mapped_Clean %>% filter(Decade == "2010"),
                   ~Longitude, 
                   ~Latitude, 
                   color = ~pal(Decade),
                   group = "2010s",
                   popup = ~paste(Internship.Site, "<br>", Location),
                   clusterOptions = markerClusterOptions()) %>% 
  addCircleMarkers(data = Tidy_Mapped_Clean %>% filter(Decade == "2020"),
                   ~Longitude, 
                   ~Latitude, 
                   color = ~pal(Decade),
                   group = "2020s",
                   popup = ~paste(Internship.Site, "<br>", Location),
                   clusterOptions = markerClusterOptions()) %>% 
  addLayersControl(overlayGroups = c("1970s", "1980s", "1990s", "2000s", "2010s", "2020s"),
                   options = layersControlOptions())
```

